### 2021-07-12 chapters/2_introduction.tex
changes

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering the mutants\cite{Ma2016,Yu2019PossibilityScope}.
Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.

To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated.
We will do this with two different approaches.
Our first approach will be a qualitative approach and second a quantitative approach. 
We structure our research on the following research questions:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement a clustering algorithm that will cluster a characteristic of the mutants. 
One or more mutants of each cluster will be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.

@@

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. 
\todo{finish this at the end}
Chapter~\ref{ch:research} describes ... 
Results are shown in Chapter~\ref{ch:results} and discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work}, contains the work related to this thesis.
Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

>>>>

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering the mutants\cite{Ma2016,Yu2019PossibilityScope}.
Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
\newline
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated.
We will do this with two different approaches.
Our first approach will be a qualitative approach and second a quantitative approach. 
We structure our research on the following research questions:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement a clustering algorithm that will cluster a characteristic of the mutants. 
We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.

@@

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. 
Chapter \ref{ch:mutation-tool} compares the different mutation testing tools to decide on a tool to use for this research.
In chapter \ref{ch:reserach_question1} we identify characteristics to represent mutants and design an experiment to cluster them.
In chapter \ref{ch:reserach_question2} we select a machine learning model and design an experiment to cluster the mutants with this model.
Results are shown in Chapters \ref{ch:results_rq1} and \ref{ch:results_rq2} for the experiments in described Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2} respectively. 
The results are then discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work}, contains the work related to this thesis.
Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

### 2021-07-12 chapters/3_background.tex
minor tweaks exp2 design + results
changes

\chapter{Background}
\label{ch:background}
This chapter will present the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
Next, we will compare existing mutation testing tools. We start of by reviewing existing literature and then extend the comparison. \todo{extend this later on?}
\section{Terminology}
\todo{do this when we finished experiments and all}
\todo{subsuming mutants}
\todo{levensteihn distance}
\todo{maven and gradle}
\section{Project selection}
We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
The unfiltered sample contains 50+ projects.
From there we filter out all the projects that contain failing tests and that aren't libraries or applications.
The result consists of a sample with sixteen projects in total.

@@

\section{Machine learning models for clustering}
\label{ch:neuralNetworkSurvey}
There are many different types of machine learning models with their respective applications.
For this thesis we will focus on machine learning models designed for clustering.
\newline
Machine learning algorithms are divided in supervised and unsupervised learning algorithms\cite{supervisedUnsupervised}.
Supervised learning is the machine learning task of learning a function that maps an input to an output based on example input-output pairs\cite{supervisedUnsupervised}.
In contrast to supervised learning, unsupervised learning shows self-organization that captures patterns as neuronal predilections or probability densities\cite{supervisedUnsupervised}.
Unsupervised learning algorithms are commonly used for classification and categorization\cite{supervisedUnsupervised}.
Unsupervised learning is a type of algorithm that learns patterns from untagged data\cite{supervisedUnsupervised}.
These machine learning algorithms are commonly used for clustering and dimensionality reduction\cite{supervisedUnsupervised}.
\newline
K.L. Du surveyed different machine learning models used for clustering\cite{Du2010Clustering:Approach}.
The models he discusses are:
\begin{itemize}
    \item Self organizing map.
    \item Learning vector quantization.
    \item C-means(K-means) clustering.
    \item Mountain clustering.
    \item Neural gas.
    \item ART networks.
    \item Fuzzy c-means.
\end{itemize}
K.L. Du describes the mathematical principles on which the machine learning algorithms are based as well as the origin, pros and cons of each model.
\newline
He also touches on subjects relevant to clustering and machine learning models such as the under-utilization problem, fuzzy clustering, robust clustering, clustering based on non- Euclidean distance measures, supervised clustering, and hierarchical clustering.
Machine learning model variants and their references are also discussed.
\newline
He closes his paper with computer simulations of some of the machine learning models he surveyed.

>>>>

\chapter{Background}
\label{ch:background}
This chapter will present the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
Next, we will compare existing mutation testing tools. We start of by reviewing existing literature and then extend the comparison. \todo{extend this later on?}
\section{Terminology}
\todo{do this when we finished experiments and all}
\todo{subsuming mutants}
\todo{levensteihn distance}
\todo{maven and gradle}
\section{Project selection}
\label{ch:project_selection}
We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
The unfiltered sample contains 50+ projects.
From there we filter out all the projects that contain failing tests and that aren't libraries or applications.
The result consists of a sample with fifteen projects in total.

@@

\section{Machine learning models for clustering}
\label{ch:neuralNetworkSurvey}
There are many different types of machine learning models with their respective applications.
For this thesis we will focus on machine learning models designed for clustering.
\newline
Machine learning algorithms are divided in supervised and unsupervised learning algorithms\cite{supervisedUnsupervised}.
Supervised learning is the machine learning task of learning a function that maps an input to an output based on example input-output pairs\cite{supervisedUnsupervised}.
In contrast to supervised learning, unsupervised learning shows self-organization that captures patterns as neuronal predilections or probability densities\cite{supervisedUnsupervised}.
Unsupervised learning algorithms are commonly used for classification and categorization\cite{supervisedUnsupervised}.
Unsupervised learning is a type of algorithm that learns patterns from untagged data\cite{supervisedUnsupervised}.
These machine learning algorithms are commonly used for clustering and dimensionality reduction\cite{supervisedUnsupervised}.
\newline
K.L. Du surveyed different machine learning models used for clustering\cite{Du2010Clustering:Approach}.
The models he discusses are:
\begin{itemize}
    \item Self organizing map.
    \item Learning vector quantization.
    \item C-means(K-means) clustering.
    \item Mountain and subtractive clustering.
    \item Neural gas.
    \item ART networks.
    \item Fuzzy c-means.
\end{itemize}
K.L. Du describes the mathematical principles on which the machine learning algorithms are based as well as the origin, pros and cons of each model.
\newline
He also touches on subjects relevant to clustering and machine learning models such as the under-utilization problem, fuzzy clustering, robust clustering, clustering based on non- Euclidean distance measures, supervised clustering, and hierarchical clustering.
Machine learning model variants and their references are also discussed.
\newline
He closes his paper with computer simulations of some of the machine learning models he surveyed.

### 2021-07-12 chapters/7_discussion.tex
changes

\chapter{Discussion}
\label{ch:discussion}
In this chapter, we discuss the results of our experiments on clustering the mutants.
First we discuss the results of the first experiment. 
We will discuss the all the variant results together and then compare the results between the variants.
Second we will discuss the results of the second experiment and compare them to those of the first.

And last we will discuss the practicality of our research.

\section{RQ 1: Identifying and clustering mutant characteristics}
The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is below the statistical threshold.
As hypothesised we observe that it is possible to cluster mutants with the set of characteristics we identified while maintaining effectiveness and reducing the amount executed.
The hypothesis holds for all the projects and variants of the experiment.
66\% of the displayed results had a deviation of less than 1\%. 
This includes both variations of the experiment.

\begin{finding}
    With a reduction up to 75\% of the total amount of mutants we can still maintain effectiveness.
\end{finding}
In both variants we observe a reduction in average accuracy when reducing the amount of clusters.
These reductions are 0.1\% and 0.04\% for clustering with and without Levenshtein distance respectively.
We can also see that the average cluster accuracy is lower when reducing the amount executed.
\begin{finding}
    The more you reduce the amount of mutants executed the more accuracy is lost.
\end{finding}
The results show that the average accuracy for reductions of 25\% and 50\% is the same.
This observation shows true for both variants of the experiment which is remarkable.
\begin{finding}
    We can reduce the amount of mutants executed in our data up to half of the original amount before reducing the accuracy even more.
\end{finding}
It is also remarkable to see the min and max accuracy of the all clusters are 50\% to 100\% respectively.
Based on these results we can state the following:
\begin{itemize}
    \item For every project there was at least one cluster that consisted only of killed or survived mutants.
    \item For every project there was at least one cluster that evenly divided between killed and survived mutants.
\end{itemize}
With the exception of Google Auto Service and Commons Text, with a reduction of 25\% and all characteristics, all projects showed an average accuracy above 90\%.
The combination of hierarchical clustering and the identified characteristics results in high accuracy clusters.
As logic dictates we expect that a higher accuracy means a score closer to that of the original.
However we do not observe this behaviour in our results.
This is due to our random sampling method.
If we would select a mutant in the majority of each cluster the accuracy per cluster would correlate to the deviation in mutation score.
Unfortunately we do not know the majority of a cluster as we do now know if a mutant survives or is killed before execution.
\begin{finding}
    Random sampling a mutant per cluster may result in a bigger deviation between original score and weighted score.
\end{finding}
As stated in Chapter \ref{ch:results_rq1} the calculation of the Levenshtein distance took a long time.
For this reason we repeated the experiment without the Levenshtein distance.
Comparing the results between these variants we can observe a max deviation of 0.05\% between average difference of the mutation score of.
In 66\% of the results displayed in Chapter \ref{ch:results_rq1} clustering without the Levenshtein distance is more accurate.
The biggest difference in mutation score measured between all the projects and their reductions is 3.63\%.
Clustering without Levenshtein distance is on average more accurate.
\begin{finding}
    While clustering without the Levenshtein distance was more accurate in our results, the Levenshtein distance characteristic does not significantly increase or decrease the cluster accuracy.
\end{finding}

\section{RQ 2: Training a machine learning model and clustering mutant characteristics}
The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is above the statistical threshold.
The biggest deviation measured is 22.39\%.
This is significantly higher than the biggest deviation measured in the first experiment.
56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
\begin{finding}
Our hypothesis does not hold for the clustering the mutants with the machine learning model trained in our experiment.
\end{finding}
We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents of the data points.
The number of clusters compared to the cluster amounts of the hierarchical clustering is significantly lower.
While we do not know the exact clusters centroids the subtractive model calculated, a low cluster amount indicates close relationships between certain mutants.
\begin{finding}
The mutants represented as characteristics in our data set per project have, according to the subtractive clustering model, close relationships to certain other mutants in the same data set.
\end{finding}
We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
This is higher than the minimum of the hierarchical clustering.
This means that in all project there was not one cluster that was evenly divided.
While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
The highest measured accuracy  measured over all variants and projects is 97.80\% with an average of 83.79\%
This means that in all project there was not one cluster that consisted of only survived or killed mutants.
\begin{finding}
The overall accuracy of the clusters generated by the machine learning model in both variants if lower than that of the hierarchical clustering.
\end{finding}
When we compare the accuracy per project between the machine learning model and hierarchical clustering we can observe that the lower accuracy of the machine learning model clusters have a higher impact on the deviation in mutation score.
This is caused by the relatively low amount of clusters.
One cluster contains more mutants which results in a bigger weight per cluster.
\begin{finding}
The inaccuracy of clusters has a bigger impact when the cluster amount is lower.
\end{finding}

@@

\section{Threats to validity}
What could affect the validity of your research? Think of pitfalls of your research method, experimental setup, interpreting the results.
- pit generated mutants -> it's hard to know if we can cluster ALL possible mutants
- Selected sources
- equivalent mutants

>>>>

\chapter{Discussion}
\label{ch:discussion}
In this chapter, we discuss the results of our experiments on clustering the mutants.
First we discuss the results of the first experiment. 
We will discuss the all the variant results together and then compare the results between the variants.
Second we will discuss the results of the second experiment and compare them to those of the first.
And last we will discuss the practicality of our research.

\section{RQ 1: Identifying and clustering mutant characteristics}
The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is below the statistical threshold.
As hypothesised we observe that it is possible to cluster mutants with the set of characteristics we identified while maintaining effectiveness and reducing the amount executed.
The hypothesis holds for all the projects and variants of the experiment.
66\% of the displayed results had a deviation of less than 1\%. 
This includes both variations of the experiment.
\newline
The results prove that the relations between the characteristics we identified can be used to cluster mutants efficiently. 
This can be explained by the correlation between the result of a mutant and the euclidean distance used to decide whether a mutants should be clustered together.
If two mutants are grouped it is \textit{n} percent likely that the mutants have the same result where \textit{n} is the cluster accuracy.
\begin{finding}
    With a reduction up to 75\% of the total amount of mutants we can still maintain effectiveness.
\end{finding}
In both variants we observe a reduction in average accuracy when reducing the amount of clusters.
These reductions are 0.1\% and 0.04\% for clustering with and without Levenshtein distance respectively.
We can also see that the average cluster accuracy is lower when reducing the amount executed.
\begin{finding}
    The more you reduce the amount of mutants executed the more accuracy is lost.
\end{finding}
The results show that the average accuracy for reductions of 25\% and 50\% is the same.
This observation shows true for both variants of the experiment which is remarkable.
\begin{finding}
    We can reduce the amount of mutants executed in our data up to half of the original amount before reducing the accuracy even more.
\end{finding}
It is also remarkable to see the min and max accuracy of the all clusters are 50\% to 100\% respectively.
Based on these results we can state the following:
\begin{itemize}
    \item For every project there was at least one cluster that consisted only of killed or survived mutants.
    \item For every project there was at least one cluster that evenly divided between killed and survived mutants.
\end{itemize}
With the exception of Google Auto Service and Commons Text, with a reduction of 25\% and all characteristics, all projects showed an average accuracy above 90\%.
The combination of hierarchical clustering and the identified characteristics results in high accuracy clusters.
As logic dictates we expect that a higher accuracy means a score closer to that of the original.
However we do not observe this behaviour in our results.
This is due to our random sampling method.
If we would select a mutant in the majority of each cluster the accuracy per cluster would correlate to the deviation in mutation score.
Unfortunately we do not know the majority of a cluster as we do now know if a mutant survives or is killed before execution.
\begin{finding}
    Random sampling a mutant per cluster may result in a bigger deviation between original score and weighted score.
\end{finding}
As stated in Chapter \ref{ch:results_rq1} the calculation of the Levenshtein distance took a long time.
For this reason we repeated the experiment without the Levenshtein distance.
Comparing the results between these variants we can observe a max deviation of 0.05\% between average difference of the mutation score of.
In 66\% of the results displayed in Chapter \ref{ch:results_rq1} clustering without the Levenshtein distance is more accurate.
The biggest difference in mutation score measured between all the projects and their reductions is 3.63\%.
Clustering without Levenshtein distance is on average more accurate.
\begin{finding}
    While clustering without the Levenshtein distance was more accurate in our results, the Levenshtein distance characteristic does not significantly increase or decrease the cluster accuracy.
\end{finding}

\section{RQ 2: Training a machine learning model and clustering mutant characteristics}
The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is above the statistical threshold.
The biggest deviation measured is 22.39\%.
This is significantly higher than the biggest deviation measured in the first experiment.
56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
\begin{finding}
Our hypothesis does not hold for the clustering the mutants with the machine learning model trained in our experiment.
\end{finding}
We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents of the data points.
The number of clusters compared to the cluster amounts of the hierarchical clustering is significantly lower.
While we do not know the exact clusters centroids the subtractive model calculated, a low cluster amount indicates close relationships between certain mutants.
\begin{finding}
The mutants represented as characteristics in our data set per project have, according to the subtractive clustering model, close relationships to certain other mutants in the same data set.
\end{finding}
We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
This is higher than the minimum of the hierarchical clustering.
This means that in all project there was not one cluster that was evenly divided.
While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
The highest accuracy measured over all variants and projects is 97.80\% with an average of 83.79\%
This means that in all project there was not one cluster that consisted of only survived or killed mutants.
\begin{finding}
The overall accuracy of the clusters generated by the machine learning model in both variants is lower compared to that of the hierarchical clustering.
\end{finding}
When we compare the accuracy per project between the machine learning model and hierarchical clustering we can observe that the lower accuracy of the machine learning model clusters have a higher impact on the deviation in mutation score.
This is caused by the relatively low amount of clusters.
One cluster contains more mutants which results in a bigger weight per cluster.
\begin{finding}
The inaccuracy of clusters has a bigger impact on mutation score when the cluster amount is lower.
\end{finding}

@@

\section{Threats to validity}
In this section, we discuss the threats to internal validity, external validity and reliability.
\subsection{Internal validity}
\subsubsection{Equivalent mutants}
Equivalent mutants are mutants that are syntactically different but functionally the same.
These mutants influence the mutation score while giving the same result as their equivalent counterpart.
In our research equivalent mutants could give a false image of the accuracy of the clusters and the overall accuracy if such a mutant was randomly selected.
PIT contains some functionality for reducing equivalent mutants\cite{pit}.
We also reduced the chance of inaccuracy by repeating our experiment.

\subsubsection{Test set representativity}
The software projects we used in our research were systematically selected.
Chapter \ref{ch:project_selection} provides an indication but not a wide representation of industrial software projects.
The conclusions drawn regarding the increase performance by reducing the mutants are based on the comparison between our own baseline measurements.
Another threat which is the result of the representation of our test set is the representation of the mutants.
While our data set contains around 1 million mutants we do not know if we have seen all mutants that exist.
This will always be a problem as it would be nearly impossible to cluster and mutation test every piece of Java code, that conforms to the requirements of PIT, ever written.

\subsubsection{Project configuration}
Some tests may not be recognized by PIT due to some external data required for the test.
Another reason may be that the test uses global variables that are not available to PIT.
We encountered some of these tests in our data set.
While these tests did pass when run with only the unit test framework they would not when run by PIT.
These projects still met the requirements stated in Chapter \ref{ch:project_selection}.
To solve this issue we removed these tests from the test suite.
By removing these specific tests we may have slightly, tough negligible, deviated from the individual real-world settings for respective projects.

\subsubsection{Expertise}
While this research is in the domain of software engineering it borrows techniques from data science.
The researcher's field of expertise is software engineering.
Since the researcher had to use techniques from different domains we can only guarantee that the research executed was done to the best of their ability.
Thus the research may contain sub optimal design choices which may have been made from a software engineering perspective.

\subsection{External validity}
We clustered mutants with two different techniques.
Our results are not generalisable to projects written in other programming languages as well as projects that cannot be run by PIT.
As a result of using PIT we limit ourselves to two testing frameworks; JUnit and TestNG.
PIT is still under active development and may expand its applicability, which in turn increases the applicability of this research.

### 2021-07-12 chapters/8_related_work.tex
changes

\chapter{Related work}
\label{ch:related_work}
We divide the related work into the following categories:  ...

>>>>

\chapter{Related work}
\label{ch:related_work}
Clustering mutants has been done before.
The research that exists tries to cluster mutants by defining the centroid of a cluster and cluster the mutants on that definition.
In our research extends mutation clustering by defining a centroid that is dynamic.
In other words the centroids of our clusters are defined by the mutants itself.

\subsection{Clustering overlapped mutants}
\label{ch:overlapping_mutants}
Ma et al., clustered mutants by overlap\cite{Ma2016}.
They defined the term overlap as as mutants that are functionally equivalent.
This is close to an equivalent mutant but not the same.
They explain that an equivalent mutants is functionally identical to the original source code while an overlapped mutant is functionally identical to at least one other mutant.
If a mutant does not have an overlapped mutant that mutant would be a cluster on its own.
The achieved a reduction, in mutants executed, of 10\%.
They write in their research that it is limited to the mutator operators on expression level. 
If a mutator did not generate two or more mutants they could not detect overlapping mutants.
Their future work includes adding more mutators and widening the scope level to statements or blocks.
\newline
The research does not include efficiency.
While they show that they can cluster mutants based on overlap they do not show how efficient the mutation testing process is compared to executing a full set of mutants.
A threat to their validity, which is not addressed, is the fact that two overlapping mutants may not be challenged by the same tests and thus can result in different results.
Our research solves this problem by taking into account the location of the mutants.
Chapter \ref{ch:characteristics_location} describes what characteristics we use.

\subsection{Clustering by scope}
Yu et al., extended the research about overlapping mutants\cite{Yu2019PossibilityScope}.
They added more mutators to the scope.
The also extended the research to include statements and blocks.
Their results shows an increase in clustered mutants.
The problems mentioned in Chapter \ref{ch:overlapping_mutants} are not addressed in this research.

\subsection{Clustering Hamming distance}
Ji et al., did a qualitative domain analysis on mutants\cite{Ji2009}.
As a result they identified the Hamming distance for to cluster with. 
The Hamming distance is, like the Levenshtein distance, a similarity measure.
They use the k-means to cluster the mutants with the Hamming distance.
They write that the reduced test set in their experiment is still as strong as the original test set\cite{Ji2009}.
Ji et al., were successful in clustering mutants with Hamming distance and the k-means algorithm.
But they do remark that their research still has problems with rationally determining the domains of their clusters.
Our research does not have this problem.
In our reserach the centre of of the domains are decided by the characteristics of the mutants.

\subsection{Spectral clustering}
Wei et al., makes use of an intelligent technique, namely spectral clustering, to improve the efficacy of mutant reduction\cite{Wei2021SpectralTesting}.
They defined multiple definitions for the mutants; similarity distance, distance between the mutants and a killing matrix of the mutants.
With these definitions they reduce the mutants based on their proposed reduction method.
This method is based on spectral clustering (SCMT), including the determination method of the number of clusters, spectral clustering of mutants, and selection of representative mutants.
The reduced were then clustered with a classical clustering algorithm.
Their results are promising and show good cluster accuracy.
They write that there is still work left to do in optimizing the matrices and clustering algorithms.
Our goal overlaps with that of Wei, the difference is that we used a different methodology to achieve that goal.

\subsection{Clustering similarity}
Hussain et al., used the k-means and agglomerative clustering algorithm to cluster the mutants according to a similarity measure. 
They used the Hamming distance a similarity measure.
They calculated the distance and used this as data te feed into the clustering algorithms.
The number of clusters and the initial position of the cluster center in the k-means algorithm are difficult to determine, and the process of the agglomerative clustering algorithm makes it difficult to correct the existing cluster formation\cite{Hussain2008}.
They did take into account efficiency.
Hussain et al, had the same goals as our research. 
The difference is that our research used different and more elaborated methods to cluster the mutants.

\subsection{Generalizing mutants}
Wilinski et al., tried to generalize the mutants by defining metrics.
Each metric is calculated for a specific mutator.
The three metrics are usefulness, frequency and dependency.
Combining the results of usefulness and frequency metrics, they observed that reducing the number of generated mutants gives noticeable mutant cost reduction without a loss of the mutation score accuracy.
Wilinski et al., their research is very narrowly scoped to the specific mutators they decided to do research on.
Our research does not contain this limitation.
The mutator is used as a characteristic in our research instead of limiting our research it helps define the the mutant.
